Place of death in NRS COVID-19 deaths: visualisation for 2023 data

### 2023-04-20 covid_nrs_place_of_death_visualisations_2023.Rmd
made changes to accommodate the disaggregatd data by age,sex,LA/HB ne longer being available alongside place of death added figure 2 for the paper

date_range_text <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2023, '%d %B %Y')} (inclusive).")

@@

only_2020_text <- glue::glue("Data from 2020 included, starting with w/c {format(earliest_date)}.")
only_2023_text <- glue::glue("Data for 2023 only include weeks 1-{format(weeks_available_2023)} (w/c {format(most_recent_date_available_2023)}).")

@@

```{r}
(fig_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2023_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)
fig_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

```{r}
fig_cause_of_death_annual_proportion_by_place <-
  current_and_historical_deaths_by_cause_and_location %>%
  filter(!cause_of_death %in% c("All")) %>%
  filter(!year %in% c("2020-2021")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home", "Other"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  mutate(
   cause_of_death = factor(cause_of_death %>% str_wrap(., width = 15), levels = order_of_causes_of_death %>% str_wrap(., width = 15)),  # needed to allow converting to a number for plotting the line segments; str_wrap breaks the text up into lines
  ) %>%
  ggplot(aes(x = year, y = proportion, colour = place_of_death, group = place_of_death)) +
  geom_line() +
  geom_point() +
  geom_text(data = ~filter(.x, year==2022), aes(x=4.2, y=proportion, label="*"), inherit.aes = FALSE) +  # 4.2 here is 4th (final) place on the scale!
  facet_rep_wrap(~cause_of_death, repeat.tick.labels = TRUE) +
  scale_colour_viridis_d() +
  scale_y_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
  labs(
    x = "Year",
    y = "Percentage yearly deaths by place (%)",
    title = "In 2020 and 2021 the share of deaths at <b style='color:#31688E;'>Home</b> increased for all causes.\nDuring the pandemic, more cancer deaths happened at <b style='color:#31688E;'>Home</b> than in <b style='color:#440154;'>Hospital</b>.",
    subtitle = paste0("\n* ",only_2023_text),
    caption = paste0(c("The 2015-2019 number of deaths is the annual mean for the period.",death_registrations_text,source_nrs_text), collapse = "\n")
  ) +
  theme(
    panel.spacing = unit(1, units = "lines"),
    legend.position = "top", legend.title = element_blank(),
    plot.title = element_textbox_simple()  # to render the html in the title
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

@@

```{r}
(fig_home_deaths_over_time_by_hb_4_week_average <-
  merged_deaths_hb %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2023)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), size = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by NHS Health Board",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text,death_registrations_text,source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

@@

```{r}
(fig_home_deaths_over_time_by_la_4_week_average <-
  merged_deaths_la %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2023)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), size = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by Scottish Local Authority",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text,death_registrations_text,source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

@@

```{r}
(fig_proportion_of_deaths_by_place_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2023_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

```{r}
(fig_proportion_of_deaths_by_place_by_la <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "la") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2023_text),
    caption = paste0(c("Local Authorities sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

```{r}
(fig_proportion_of_deaths_at_home_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(merged_proportions_of_deaths_by_place %>% select(year, place_of_death, proportion_scotland = proportion_of_total), by = c("year","place_of_death")) %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(proportion_of_total, proportion_scotland), names_to = "measure", values_to = "proportion_of_total") %>%
  mutate(measure = if_else(measure == "proportion_of_total", true = "Health board", false = "Scotland")) %>%
  ggplot(., aes(x=year, y=proportion_of_total, linetype = measure)) +
  geom_point() +
  geom_line() +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  labs(
    linetype = NULL,
    x = NULL, y= "Proportion of yearly deaths that happened at home or other non-institution (%)",
    subtitle = paste0("* ",only_2023_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

>>>>

date_range_text <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2023, '%d %B %Y')} (inclusive).")
date_range_text_disaggregated <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2023_disaggregated, '%d %B %Y')} (inclusive).")

@@

only_2020_text <- glue::glue("Data from 2020 included, starting with w/c {format(earliest_date)}.")

only_2023_text <- glue::glue("Data for 2023 only include weeks 1-{format(weeks_available_2023)} (w/c {format(most_recent_date_available_2023)}).")
only_2023_disaggregated_text <- glue::glue("Data for 2023 only include weeks 1-{format(weeks_available_2023_disaggregated)} (w/c {format(most_recent_date_available_2023_disaggregated)}).")

@@

```{r}
(fig_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(linewidth = 1) +
  geom_text(data = ~filter(.x, year==latest_year), aes(x=latest_year+0.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2023_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)
fig_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

```{r}
fig_cause_of_death_annual_proportion_by_place <-
  current_and_historical_deaths_by_cause_and_location %>%
  filter(!cause_of_death %in% c("All")) %>%
  filter(!year %in% c("2020-2021")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home", "Other"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  mutate(
   cause_of_death = factor(cause_of_death %>% str_wrap(., width = 15), levels = order_of_causes_of_death %>% str_wrap(., width = 15)),  # needed to allow converting to a number for plotting the line segments; str_wrap breaks the text up into lines
  ) %>%
  ggplot(aes(x = year, y = proportion, colour = place_of_death, group = place_of_death)) +
  geom_line() +
  geom_point() +
  geom_text(data = ~filter(.x, year==as.character(latest_year)), aes(x=n_distinct(current_and_historical_deaths_by_cause_and_location$year)+0.2, y=proportion, label="*"), inherit.aes = FALSE) +  # cobble together the final year position from the data!
  facet_rep_wrap(~cause_of_death, repeat.tick.labels = TRUE) +
  scale_colour_viridis_d() +
  scale_y_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
  labs(
    x = "Year",
    y = "Percentage yearly deaths by place (%)",
    title = "In 2020 and 2021 the share of deaths at <b style='color:#31688E;'>Home</b> increased for all causes.\nDuring the pandemic, more cancer deaths happened at <b style='color:#31688E;'>Home</b> than in <b style='color:#440154;'>Hospital</b>.",
    subtitle = paste0("\n* ",only_2023_text),
    caption = paste0(c("The 2015-2019 number of deaths is the annual mean for the period.",death_registrations_text,source_nrs_text), collapse = "\n")
  ) +
  theme(
    panel.spacing = unit(1, units = "lines"),
    legend.position = "top", legend.title = element_blank(),
    plot.title = element_textbox_simple()  # to render the html in the title
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

@@

```{r}
(fig_home_deaths_over_time_by_hb_4_week_average <-
  merged_deaths_hb %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2023)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), linewidth = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by NHS Health Board",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text,death_registrations_text,source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

@@

```{r}
(fig_home_deaths_over_time_by_la_4_week_average <-
  merged_deaths_la %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2023)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), linewidth = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by Scottish Local Authority",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text,death_registrations_text,source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

@@

```{r}
(fig_proportion_of_deaths_by_place_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(linewidth = place_of_death)) +
  geom_text(data = ~filter(.x, year==latest_year), aes(x=latest_year+0.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_linewidth_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2023_disaggregated_text),
    caption = paste0(c("Health boards ordered by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

```{r}
(fig_proportion_of_deaths_by_place_by_la <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "la") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(linewidth = place_of_death)) +
  geom_text(data = ~filter(.x, year==latest_year), aes(x=latest_year+0.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_linewidth_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2023_disaggregated_text),
    caption = paste0(c("Local Authorities ordered by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

```{r}
(fig_proportion_of_deaths_at_home_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(merged_proportions_of_deaths_by_place %>% select(year, place_of_death, proportion_scotland = proportion_of_total), by = c("year","place_of_death")) %>%
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(proportion_of_total, proportion_scotland), names_to = "measure", values_to = "proportion_of_total") %>%
  mutate(measure = if_else(measure == "proportion_of_total", true = "Health board", false = "Scotland")) %>%
  ggplot(., aes(x=year, y=proportion_of_total, linetype = measure)) +
  geom_point() +
  geom_line() +
  geom_text(data = ~filter(.x, year==latest_year), aes(x=latest_year+0.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_linewidth_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  labs(
    linetype = NULL,
    x = NULL, y= "Proportion of yearly deaths that happened at home or other non-institution (%)",
    subtitle = paste0("* ",only_2023_disaggregated_text),
    caption = paste0(c("Health boards ordered by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

## Absolute num. home deaths by LA

```{r}
# TODO: remove theme_minimal when done testing
# TODO: add Scotland graph also
# TODO: check  if all PoD would work?

la_group <- tibble(  # group LAs into 4 groups of 8
  ref_area = factor(order_la_by_population),
  # la_group = rep(1:4, each = 8))
  la_group = rep(1:6, times = c(6,5,5,5,5,6))
  ) %>%
  group_by(la_group) %>%
  mutate(la_group_label = paste0(ref_area[1], " - ", ref_area[n()])) %>%
  ungroup %>%
  mutate(la_group_label = factor(la_group_label, levels = unique(la_group_label)))
